---
title: "MOG RTX meta analysis"
author: "Bo Sun"
date: "2022-03-30"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, warning=FALSE, message=FALSE}
library(meta)
library(metafor)
library(ggrepel)
library(dmetar)
library(smd)
library(tidyverse)
library(PerformanceAnalytics)
library(ggplot2); theme_set(theme_bw(base_size = 15)+
                                    theme(axis.text.x = element_text(angle = 45, hjust=1), 
                                          strip.text = element_text(colour = 'black', face="bold",size=12), 
                                          panel.grid.major = element_blank(), 
                                          panel.grid.minor = element_blank(),
                                          panel.border = element_rect(size = 0.7),
                                          axis.ticks.length=unit(.10, "cm"),
                                          axis.ticks = element_line(size=0.7),
                                          strip.background = element_blank()))
```

##Loading data and prepping tables
It is worth noting that using MD results in higher heterogeneity values than using SMD. Here I use MD as we have discussed. The significance does not change, but it affects the multi-model inference later. Using SMD, there is no significant influence... using MD there is. 
```{r, warning=FALSE, message=FALSE, echo=FALSE}
### load data
curdir = "/Users/bosun/Documents/DPhil/OANG/6\ -\ Collaborations/2021-12-27_Val_meta/"
mog <- read.csv("/Users/bosun/Documents/DPhil/OANG/6\ -\ Collaborations/2021-12-27_Val_meta/data/2022-03-27-MOG_ARR_ALL.csv")
mogedss <- read.csv("/Users/bosun/Documents/DPhil/OANG/6\ -\ Collaborations/2021-12-27_Val_meta/data/2022-03-27-MOGEDSS.csv")
mogedss <- mogedss[complete.cases(mogedss), ]

##remove Shu paper 
mog <- mog[!grepl("Shu", mog$STUDY),]
mogedss <- mogedss[!grepl("Shu", mogedss$STUDY),]

#mog <- mog[-11,]
mogmeta <- escalc(measure="MD", m1i=mog$mean_ARR_postRTX, m2i=mog$mean_ARR_preRTX, 
       sd1i=mog$SD.2, sd2i=mog$SD.1, n1i=mog$n, n2i=mog$n, data=mog)
mogmeta$sei<-sqrt(mogmeta$vi)
mogmeta$Group <- "MOG"

mogedssmeta <- escalc(measure="MD", m1i=mogedss$mean_EDSS_postRTX, m2i=mogedss$mean_EDSS_preRTX, 
       sd1i=mogedss$SD.2, sd2i=mogedss$SD.1, n1i=mogedss$n, n2i=mogedss$n, data=mogedss)
mogedssmeta$sei<-sqrt(mogedssmeta$vi)
mogedssmeta$Group <- "MOG"

##load nmosd data and merge with mog
nmosd <- read.csv("/Users/bosun/Documents/DPhil/OANG/6\ -\ Collaborations/2021-12-27_Val_meta/data/2022-03-27-NMO_ARR_ALL.csv")
nmosdedss <- read.csv("/Users/bosun/Documents/DPhil/OANG/6\ -\ Collaborations/2021-12-27_Val_meta/data/2022-03-27-NMO_EDSS_ALL.csv")

nmosdmeta <- escalc(measure="MD", m1i=nmosd$mean_ARR_postRTX, m2i=nmosd$mean_ARR_preRTX, 
                  sd1i=nmosd$SD.2, sd2i=nmosd$SD.1, n1i=nmosd$n, n2i=nmosd$n, data=nmosd)
nmosdmeta$sei<-sqrt(nmosdmeta$vi)
nmosdmeta$Group <- "NMOSD"

write.csv(nmosdmeta, "/Users/bosun/Documents/DPhil/OANG/6\ -\ Collaborations/2021-12-27_Val_meta/data/nmosdmeta.csv")

nmosdedssmeta <- escalc(measure="MD", m1i=nmosdedss$mean_EDSS_postRTX, m2i=nmosdedss$mean_EDSS_preRTX, 
       sd1i=nmosdedss$SD.2, sd2i=nmosdedss$SD.1, n1i=nmosdedss$n, n2i=nmosdedss$n, data=nmosdedss)
nmosdedssmeta$sei<-sqrt(nmosdedssmeta$vi)
nmosdedssmeta$Group <- "NMOSD"

mergemeta <- rbind(mogmeta, nmosdmeta)
mergeedssmeta <- rbind(mogedssmeta, nmosdedssmeta)

write.csv(mergemeta, "data/mergemeta.csv")
write.csv(mergemeta, "data/mergeedssmeta.csv")

mergemeta <- read.csv("data/mergemeta.csv")
mergemeta <- read.csv("data/mergeedssmeta.csv")

mogmetagen <- metagen(TE = yi,
        seTE = sei,
        studlab = STUDY,
        data = mogmeta,
        sm = "MD",
        fixed = FALSE,
        random = TRUE,
        method.tau = "REML",
        hakn = FALSE,
        title = "Mogmeta")

nmosdmetagen <- metagen(TE = yi,
        seTE = sei,
        studlab = STUDY,
        data = nmosdmeta,
        sm = "MD",
        fixed = FALSE,
        random = TRUE,
        method.tau = "REML",
        hakn = FALSE,
        title = "NMOSDmeta")

mogedssmetagen <- metagen(TE = yi,
        seTE = sei,
        studlab = STUDY,
        data = mogedssmeta,
        sm = "MD",
        fixed = FALSE,
        random = TRUE,
        method.tau = "REML",
        hakn = FALSE,
        title = "MogEDSSmeta")

nmosdedssmetagen <- metagen(TE = yi,
        seTE = sei,
        studlab = STUDY,
        data = nmosdedssmeta,
        sm = "MD",
        fixed = FALSE,
        random = TRUE,
        method.tau = "REML",
        hakn = FALSE,
        title = "NMOSDEDSSmeta")

mergemetagen <- metagen(TE = yi,
                      seTE = sei,
                      studlab = STUDY,
                      data = mergemeta[!is.na(mergemeta$yi),],
                      sm = "MD",
                      fixed = FALSE,
                      random = TRUE,
                      method.tau = "REML",
                      hakn = FALSE,
                      title = "mergemeta")

mergeedssmetagen <- metagen(TE = yi,
                      seTE = sei,
                      studlab = STUDY,
                      data = mergeedssmeta[!is.na(mergeedssmeta$yi),],
                      sm = "MD",
                      fixed = FALSE,
                      random = TRUE,
                      method.tau = "REML",
                      hakn = FALSE,
                      title = "mergeEDSSmeta")
```

## Assessing bias
Here we see that MOG papers are biased. We will have to state this as a limitation
```{r}
# Generate funnel plot (we do not include study labels here)
funnel.meta(mogmetagen)
title("MOG ARR")


# Plot over points

eggers.test(mogmetagen)

```

NMOSD papers are much more even. 
```{r}
# Generate funnel plot (we do not include study labels here)
funnel.meta(nmosdmetagen)
title("NMOSD ARR")


# Plot over points

eggers.test(nmosdmetagen)

```

MOG EDSS scores are not biased across studies. 
```{r}
# Generate funnel plot (we do not include study labels here)
funnel.meta(mogedssmetagen)
title("MOG EDSS")


# Plot over points

eggers.test(mogedssmetagen)
```


NMOSD EDSS scores are also not biased. 
```{r}
# Generate funnel plot (we do not include study labels here)
funnel.meta(nmosdedssmetagen)
title("NMOSD EDSS")


# Plot over points

eggers.test(nmosdedssmetagen)
```

## Forest plot MOG ARR

Here a random effects model reveals significant difference in mean ARR with lower ARRs post RTX.
```{r,fig.height=7, fig.width=10, warning=FALSE, message=FALSE}
##plot mog ARR forest plot
forest.meta(mogmetagen, 
            sortvar = yi,
            predict = TRUE, 
            print.tau2 = FALSE,
            leftlabs = c("Study", "g", "SE"))

print(mogmetagen)
```

## Forest plot NMOSD ARR
As shown before in Damata et al., NMOSD post RTX ARRs are significantly lower. 
```{r,fig.height=7, fig.width=10, warning=FALSE, message=FALSE}
##plot mog ARR forest plot
forest.meta(nmosdmetagen, 
            sortvar = yi,
            predict = TRUE, 
            print.tau2 = FALSE,
            leftlabs = c("Study", "g", "SE"))

print(nmosdmetagen)
```

## Merged ARR analysis
The merge analysis reveals that NMOSD patients benefit more than MOG patients. Between groups Q=9.09 p=0.003. 

```{r,fig.height=16, fig.width=11, warning=FALSE, message=FALSE}
##plot mog and nmosd together
mergemetagen <- update.meta(mergemetagen, 
            subgroup = Group, 
            tau.common = FALSE)
forest.meta(mergemetagen, 
            sortvar = yi,
            predict = TRUE, 
            print.tau2 = FALSE,
            subgroup = TRUE,
            leftlabs = c("Study", "MD", "SE"))
print(mergemetagen)

```

## Forest plot MOG EDSS
EDSS appeared to be significantly better in those post RTX. 
```{r,fig.height=7, fig.width=10, warning=FALSE, message=FALSE}
##plot MOG EDSS
forest.meta(mogedssmetagen, 
            sortvar = yi,
            predict = TRUE, 
            print.tau2 = FALSE,
            leftlabs = c("Study", "g", "SE"))
print(mogedssmetagen)

```

## Forest plot NMOSD EDSS

EDSS is significantly improved post RTX in NMOSD patients.
```{r,fig.height=7, fig.width=10, warning=FALSE, message=FALSE}
##plot MOG EDSS
forest.meta(nmosdedssmetagen, 
            sortvar = yi,
            predict = TRUE, 
            print.tau2 = FALSE,
            leftlabs = c("Study", "g", "SE"))
print(nmosdedssmetagen)

```

## Combined analysis of EDSS

Here we see that there is no difference between EDSS scores of patients pre and post RTX between MOG and NMOSD (p=0.4254)
```{r,fig.height=16, fig.width=11, warning=FALSE, message=FALSE}
##plot mog and nmosd together
mergeedssmetagen <- update.meta(mergeedssmetagen, 
            subgroup = Group, 
            tau.common = FALSE)
forest.meta(mergeedssmetagen, 
            sortvar = yi,
            predict = TRUE, 
            print.tau2 = FALSE,
            subgroup = TRUE,
            leftlabs = c("Study", "MD", "SE"))
print(mergeedssmetagen)

```


## Multiple regression - Univariate models
Here we removed mean baseline EDSS and prop children. As they significantly correlate with other predictors. 
```{r, warning=FALSE, message=FALSE}
##meta-regression
mogreg<- read.csv("/Users/bosun/Documents/DPhil/OANG/6\ -\ Collaborations/2021-12-27_Val_meta/data/2022-03-27-mog_meta-regression data_last.csv")

mogreg <- mogreg %>%
        dplyr::left_join(mogmeta, by = c("STUDY", "n"))

mogreg$weights <- 1/sqrt(mogreg$vi)

plot_metareg <- function(data = mogreg, variable = NULL){
        mogreg2 <- mogreg[!is.na(mogreg[[paste0(variable)]]),]
        
        reg <- rma(yi = yi,
                   vi = vi,
                   mods = mogreg2[[paste0(variable)]],
                   data = mogreg2)
        
        # Specify basic plot, mapping sex to the x-axis, effect size 'd' to the y-axis,
        # and 'weights' to the weight parameter.
        preds <- predict.rma(reg)
        plotdb <- cbind(mogreg2, preds)
        ggplot(plotdb, aes(x = plotdb[[paste0(variable)]], y = plotdb[["yi"]], label=plotdb[["STUDY"]])) +
                geom_ribbon(aes(ymin = ci.lb, ymax = ci.ub), alpha=0.05)+
                geom_line(data = plotdb,aes(x = plotdb[[paste0(variable)]],y = plotdb$pred, size = 0.5), linetype="dashed", alpha=0.1)+
        geom_text_repel(min.segment.length = 0, seed = 42, box.padding = 0.5, size=3, force=1)+
                geom_point(aes(size=weights),shape = 21, alpha=1, fill="white") +
                theme(legend.position = "none")+
                xlab(variable)+
                ylab("MD")
}
#Get modifiers
plotparam<- c("mean_time_to_RTX_months",
              "mean_onset_age", 
              "mean_fup_months",
              "mean_baseline_ARR",
              "mean_baseline_EDSS",
              "prop_relapsing",
              "prop_children",
              "prop_female")

#filter modifiers for highly correlated
mogreg[,unlist(plotparam)] %>%
        chart.Correlation()

##remove prop_children as correlated with mean age onset
## remove mean EDSS as correlated highly with age of onset
plotparam <-plotparam[!plotparam %in% c("mean_baseline_EDSS", "prop_children")]


reg_plotlist <- lapply(c(plotparam,"n"), function(x){
 p <- plot_metareg(data=mogreg, variable = x)
 return(p)
})
```


```{r,fig.height=8, fig.width=12, warning=FALSE, message=FALSE}
##plot bubble plots
patchwork::wrap_plots(reg_plotlist)
```

```{r, warning=FALSE, message=FALSE}
##get univariate model metrics
get_unimodels <- function(data, mod){
        m_reg <- rma.uni(yi = yi,
                         vi = vi,
                         mods = data[[paste0(mod)]],
                         data = data)
        
        metrics <- c(mod = mod, tau2 = round(m_reg$tau2,3),
                     tau =round(sqrt(m_reg$tau2), 3),
                     I2 = round(m_reg$I2, 3),
                     R2 = round(m_reg$R2, 3),
                     beta = round(m_reg$beta["mods",],3),
                     zval = round(m_reg$zval[2],3),
                     pval = round(m_reg$pval[2],7),
                     ci.lb = round(m_reg$ci.lb[2],3),
                     ci.ub = round(m_reg$ci.ub[2],3))
}


modregs <- lapply(c(plotparam,"n"), function(x){
        get_unimodels(mogreg, x)
})
modregs <- arrange(as.data.frame(do.call(rbind, modregs)), desc(as.numeric(R2)))

kableExtra::kable(modregs)%>% kableExtra::kable_material(c("striped", "hover"))

```

## Multi-model inference
Here,  all possible predictor combinations with multi-model inference, we see that mean ARR is the most important predictor of RTX response. 
```{r, warning=FALSE, message=FALSE}
multimodel.inference(TE = "yi", 
                     seTE = "sei",
                     data = mogreg,
                     predictors = c(plotparam),
                     interaction = TRUE)
```


## Subgroup extraction and analysis

```{r, echo=FALSE, warning=FALSE, message=FALSE}
relapsing_mog <- read.csv(paste0(curdir, "data/2022-03-27-MOG_ARR_RELAPSING.csv"))
adult_mog <- read.csv(paste0(curdir,"data/2022-03-27-MOG_ARR_ADULTS.csv"))
children_mog <- read.csv(paste0(curdir,"data/2022-03-27-MOG_ARR_CHIL.csv"))
fu12_mog <- read.csv(paste0(curdir,"data/2022-03-27-MOG_ARR_FU12M.csv"))
firstline_mog <- read.csv(paste0(curdir,"data/2022-03-27-MOG_ARR_FIRST.csv"))

subgroups <- list(relapsing_mog,adult_mog, children_mog, fu12_mog, firstline_mog)
names(subgroups) <- c("relapsing_mog","adult_mog", "children_mog", "fu12_mog", "firstline_mog")

subgroups <- lapply(subgroups, function(x){
  y <- escalc(measure="MD", m1i=x[["mean_ARR_postRTX"]], m2i=x[["mean_ARR_preRTX"]], 
       sd1i=x[["SD.2"]], sd2i=x[["SD.1"]], n1i=x[["n"]], n2i=x[["n"]], data=x)
y$sei<-sqrt(y$vi)
return(y)
})

subgroupsmetagen <- lapply(names(subgroups), function(x){
  metagen(TE = yi,
        seTE = sei,
        studlab = STUDY,
        data = subgroups[[paste0(x)]],
        sm = "MD",
        fixed = FALSE,
        random = TRUE,
        method.tau = "REML",
        hakn = FALSE,
        title = paste0(x))
})

names(subgroupsmetagen)<- c("relapsing_mog","adult_mog", "children_mog", "fu12_mog", "first_mog")
print(subgroupsmetagen)
```

## Forest plots of subgroups

### 
```{r,fig.height=9, fig.width=10}


for (i in 1:length(subgroupsmetagen)){
  forest.meta(subgroupsmetagen[[paste0(names(subgroupsmetagen[i]))]], 
            sortvar = yi,
            predict = TRUE, 
            print.tau2 = FALSE,
            leftlabs = c("Study", "g", "SE"))
  grid::grid.text(paste0(names(subgroupsmetagen[i])), .5, .9, gp=grid::gpar(cex=2))
}

```
















